---
title: "Claire's Exam 2"
format: html
---

## Missing Data  

| Types | Description |
|-------|-------------|
| MCAR | The probability of missing data for a variable is the same for all cases. Implies that causes of the missing data are unrelated to the data. |
| MAR  | The probability of missing data is related to observed variables but unrelated to unobserved information. |
| MNAR | The probability of missing data is related to unobserved variables (and probably observed variables too). |

* Identifying Missing Data: is.na(), skim(), vis_miss()
* Handling Missing Data: drop_na(), replace_na()

| Explicit missing data | Implicit missing data |
|-----------------------|-----------------------|
|Data that is explicitly marked as missing (NA) | Data that is missing but not explicitly marked as such. |

 *Use the function complete() to create that new row and plug in values of NA for the missing data.*
 
## Functions  
```{r eval=FALSE}
# Defining a function
function_name <- function(input_name_1, input_name_2 = default_value_2){

  # function body - code that does something

  return(output) 
}

# Calling a function (all valid ways)
function_name(input_name_1 = 2, input_name_2 = 4)
function_name(2, 4)
function_name(input_name_2 = 4, input_name_1 = 2)
function_name(2)
```

Example for if-else condition: 
```{r eval = FALSE}
convert_temp <- function(temp, unit) {
    if (unit=="F") {
        (temp - 32) * 5/9
    } else if (unit=="C") {
        (temp * 9/5) + 32
    }
}

convert_temp(0, unit = "C")
convert_temp(32, unit = "F")
```

## Base R 
### Vectors 
```{r eval = FALSE}
num_vec <- vector("numeric", length = 2) #empty vector: Zeros
log_vec <- vector("logical", length = 3) #empty vector: FALSE
chr_vec <- vector("character", length = 4) #empty vector: empty strings
named_vec <- c('name1' = 1, 'name2' = 2) # Named numeric vector
```
### Lists 
```{r eval = FALSE}
ex_list <- list(a = 1:3, b = c("a", "b", "c"), c = matrix(1:6, nrow = 2))
```
### Array, Matrices, Data Frame
```{r eval=FALSE}
ary <- array(NA, dim = c(2,3,4))
m <- matrix(NA, nrow = 2, ncol = 3)
mod_df <- tibble(x = 1:10, y = 1:10 + rnorm(10))
```
Subset with [], and single elements can be subsetted with $.

## Loops+Iter 
### Iteration
```{r eval = FALSE}
# seq_len() iterates over the rows of a data frame, seq_along() would iterate over columns 
for (i in seq_along(groups)) {
    print(groups[i])
}

for (i in seq_len(length(groups))) {
    print(groups[i])
}
```
### Memory Allocation (Purrr, map)
```{r eval=FALSE}
# Example 1: No storage container; append new value to existing vector
x <- integer()
bench::mark(
  for (i in 1:1e5) {
    x <- c(x, i)
  }
)

# Example 2: Using a storage container!
#Notice median (time to run), mem_alloc (how much memory you are using), and gc (garbage collection)
x <- vector('integer', length = 1e5)
bench::mark(
  for (i in 1:1e5) {
    x[i] <- i
  }
)

# Example 3: Vectorized function
#Notice median (time to run), mem_alloc (how much memory you are using), and gc (garbage collection)
bench::mark(
  x <- seq(1, 1e5, by = 1)
)

# Example 4: purrr::map()
results <- bench::mark(purrr::map_int(1:1e5, ~.x))
```

* map() returns a list
* map_chr() returns a character vector
* map_lgl() returns a logical vector
* map_int() returns an integer vector
* map_dbl() returns a numeric vector
* map_vec() returns a vector of a different (non-atomic) type (like dates)

### map2 or pmap
For iterating across rows of a data frame!
```{r eval=FALSE}
# save randomly generated data in tibble with simulation parameters
sim_data <- args |>
  mutate(rand_data = pmap(args, rnorm))
# specify variations in some arguments but leave some arguments constant across the iterations
pmap_chr(string_data, str_replace_all, string = "ppp nnn hhh")
```

### walk
No need to save an output for each iteration...
```{r eval=FALSE}
# Iterate over paths and plots to save the export images
walk2( 
  by_clarity$path,
  by_clarity$plot,
  function(path, plot){ ggsave(path, plot, width = 6, height = 6)}
)
```

## APIs 
### URLs 
| Base URL | Scheme      | Host Name               | File Path          | Query String |
|----------|-------------|-------------------------|--------------------|--------------|
| https://api.census.gov | http:// | api.census.gov | data/2019/acs/acs1 | ?get = NAME, B02015_009E, B02015_009M &for=state:* |

### Wrapper Packages 
```{r eval=FALSE}
tidycensus::get_acs(
    year = 2020,
    state = "MN",
    geography = "tract",
    variables = c("B01003_001", "B19013_001"),
    output = "wide",
    geometry = TRUE
)
```

### Example 
```{r eval=FALSE}
req <- request("https://boardgamegeek.com/xmlapi2") |> 
    req_url_path_append("search") |> 
    req_url_query(query = I("mystery+curse"), type = I("boardgame,boardgameaccessory,boardgameexpansion"))
```

Then, req_perform(), to see Content-Type that the format of the response is something called XML. We can navigate to the request URL to see the structure of this output. Use resp_body_xml() to read in the XML as an R object.
* xml_find_all(), xml_attr() 
* resp_body_json(simplifyVector = TRUE)

## Web Scraping 
*robots.txt* find what can be scraped
```{r eval=FALSE}
nih <- read_html("https://www.nih.gov/news-events/news-releases")
# Retrieve and inspect article titles
article_titles <- nih |>
    html_elements(".thumbnail-teaser__heading") |>
    html_text()
```

## Database and SQL 
1. DBI::dbConnect(duckdb::duckdb())
2. dbWriteTable(con, "flights", nycflights13::flights)
3. flights <- tbl(con, "flights") 

### with *dplyr*
* show_query()

### Composing Query 
* SELECT tells SQL that a query is coming.
* TOP(100) only returns the first 100 rows.
* Id, Title, Score, Body, Tags determines what columns are included in the result
* FROM Posts determines the source dataset. 
* ORDER BY ??? DESC syntax is similar to R’s arrange()
* GROUP BY
* JOIN keyword must go in between the two tables we want to join. Each table must be named. 

## Effective and Interactive Visualization

* Motivation & Context
* Orientation (axes, color legend, etc.)
* Highlights (one or two interesting data points)
* Big Picture (overall takeaways)

### Plotly and Shiny (Ui and Server, App)

### Code Quality Checklist
- [ ] Readability
- [ ] Meaningful Names
- [ ] Testing / Test Cases
- [ ] Efficiency
- [ ] Reproducibility
- [ ] Simple Tasks in Functions


### Data Quality Checklist 
- [ ] Data Parsing (lubridate, stringr)
- [ ] Data Joining (correct joins ie. list_rbind(), list_cbind())



